---
title: "kumar_ak1648_final_project"
author: "Akshat Kumar"
date: "5/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(lubridate)
library(MASS)
library(caret)
library(tidytext)
library(tidyr)
library(wordcloud)
library(ggplot2)
library(tidyverse)
library(NLP)
library(tm)
library(plotly)
```

Load The NYC Airbnb Open dataset which contains 16 columns and 48895 unique values 
and the full data can be obtained from http://insideairbnb.com/

I wanted to take the Airbnb from two different sources that is one the listing dataset 
and the other one is reviews datasetbut I got the dataset which had a combination 
for both of them and required subsequent amount of cleaning.

```{r }
airbnb <- read.csv("Airbnb_Listings_NYC.csv")
head(airbnb)
```

First I have introduced three important variables into our problem which is really important for our case. 
These are
• Price Group - KPI for classifying based on price
• Usage group - Listings use based on the number of reviews
• Total Revenue - product of price, reviews and nights and the decision class for our problem

```{r }
airbnb <- airbnb %>% mutate(price_group=ifelse(price < 61, "Very Low", ifelse(price < 91, "Low", ifelse(price < 131, "Moderate",
          ifelse(price < 201, "High", "Very High"))))) %>% mutate(usage_group=ifelse(number_of_reviews < 1, "None",ifelse(number_of_reviews < 4, "Low", ifelse(number_of_reviews < 10, "Moderate", ifelse(number_of_reviews < 34, "High", "Very High"))))) %>% mutate(total_net_revenue = price * number_of_reviews * minimum_nights)

head(airbnb)
```

Then in order to get a clean corpus for further analysis we need to clean or mine the reviews. 
So we will do the following steps for that remove unwanted characters, convert all words to lowercase, 
filter unnecessary large sentences, remove stopwords that neither add positive or negative impact

Then comes the final preparation step to convert into term document matrix
for Exploratory Data Anaysis step

```{r }
dfa <- data.frame(doc_id = 1:48895, text = airbnb$name, stringsAsFactors = FALSE)
clean <- Corpus(DataframeSource(dfa))

clean <- tm_map(clean, removePunctuation)
clean <- tm_map(clean, tolower)
clean <- tm_map(clean, removeNumbers)
clean <- tm_map(clean, removeWords, stopwords("english"))

dtma <-DocumentTermMatrix(clean,control = list(weighting = function(x) weightTfIdf(x, normalize = FALSE)))

wc_corpus <- sort(colSums(as.matrix(dtma)), decreasing=TRUE) 
```

Manhattan contains the most amount of listings; additionally, it looks like it is the most expensive. 
Next,Brooklyn appears to be next popular; however,with a more reasonable distribution of price listings. 
High - Very High seem to only take up 20% of the population. 
Queens has only 5000 listings, and appears very cheap. Lastly, 
the Bronx and State Island do not appear very popular for Airbnb.

```{r }
ggplot(airbnb, aes(neighbourhood_group)) + geom_bar(aes(fill = price_group)) + ggtitle("Listings Count by Borough")
```

We can see from the count of listings by room type graph that entire home/apt have high price range 
and shared have generally low range for all the price groups with private in the middle.

```{r }
ggplot(airbnb, aes(room_type)) + geom_bar(aes(fill = price_group)) + ggtitle("Count of Listings by Room Type")
```

Now we can focus on the density of the listings with respect to latitude and longitude 
and see the distribution by borough and price group. 
We can see that Manhattan, Bronx and Brooklyn are highly populated in comparison to the other boroughs. 
Price group density shows the high price density listings with the distribution of low price.

```{r }
ggplot(airbnb, aes(longitude,latitude)) + geom_density2d(aes(color = neighbourhood_group)) + ggtitle("Density of Listings by Borough")
```

```{r }
ggplot(subset(airbnb, price_group == "Very High" | price_group == "Very Low"), aes(longitude,latitude)) + geom_density2d(aes(color = price_group)) + ggtitle("Density of Listings by Price")
```

For customer centric companies like Airbnb, word clouds can help them find customers’ pain points. 
If they collect feedback from your customers, they can generate a word cloud using customers’ 
language to help identify what is most important to them. Imagine if “long wait time” cropped up
as major emphasis words in customer feedback. That should ring a warning bell.
Luxury, Loft, and Village pop out for high priced listings.

```{r }
wordcloud(names(wc_corpus), wc_corpus, min.freq = 825, max.words=100, rot.per=0.2,scale=c(2.4,1))
airbnb <- airbnb[1:1000,]
```

We can see from the population density by price graph that listing with lower price ranges 
have the most dense area and that density is inversely proportional to the price group of the listings.


```{r }
airbnb <- airbnb[, !duplicated(colnames(airbnb))]
ggplot(airbnb, aes(price)) + geom_density(aes(fill = price_group), alpha=0.35) + ggtitle("Count of Listings by Room Type")

```


```{r }
ggplot(subset(airbnb, price < 500), aes(price)) + geom_density(aes(fill = price_group), alpha=0.35) + ggtitle("Listing Population Density by Price")
```

From the below graph of population densit by number of reviews, 
it also looks like 7% of our population has 0 reviews.
Also we can observe that less dense populated areas have large reviews. 
We can observe the same behaviour by looking a the plot of population density 
with minimum nights required and both these graphs show the inverse relationship with population density. 
The first population density vs minimum nights graph is really skewed 
so we will adjust it putting a filter of less than 50 nights required to get a more reasonable graph as below.

```{r }
ggplot(airbnb, aes(number_of_reviews)) + geom_density() + ggtitle("Listing Population Density by Number of Reviews")
```

```{r }
ggplot(subset(airbnb, minimum_nights < 50), aes(minimum_nights)) + geom_density() + ggtitle("Listing Population Density by Minimum Nights Required")
```

```{r }
filter_airbnb <- subset(airbnb, total_net_revenue > 0)

summary(filter_airbnb$total_net_revenue)
head(filter_airbnb)
```

Now we will observe the effects of of room type on the net revenue through the boxplot. 
As we can see the revenue generally comes from Entire home/apt in all the cities because travellers 
usually prefer that for their stays but we can see one exception in the form of Bronx 
in which private takes the major chunk of the revenue.

```{r }

table(factor(filter_airbnb$room_type),filter_airbnb$neighbourhood_group)

ggplot(filter_airbnb, aes(neighbourhood_group, log(total_net_revenue))) + geom_boxplot(aes(fill = factor(`room_type`))) + ggtitle("Room Type Effects on Net Revenue")

```

Now boxplot will tell us the relationship between price group and the total revenue generated by the listings. 
So there are a few things that we can analyze and these can’t be generalized in one category. 
We see in Brooklyn and Manhattan high priced group listings bring out the major revenue for these cities.
Staten Island’s listings income come from very cheap priced listing maybe 
because they don’t have that many high priced listings or people usually can’t afford. 
In Bronc less than average listing account for increase in revenue. 
So we see that different cities have different modes of revenue generation.

```{r }

table(factor(filter_airbnb$price_group),filter_airbnb$neighbourhood_group)

ggplot(filter_airbnb, aes(neighbourhood_group, log(total_net_revenue))) + geom_boxplot(aes(fill = factor(price_group))) + ggtitle("Price Group Effects on Net Revenue")

```

The model evaluates the host success and measures the financial performance of a listing 
by taking into account several factors such as location, price, reviews, availability besides other factors. 
We will predict for the two most popular destinations in New York that is
Brooklyn and Manhattan.
We will decide the dataset 80/20 for training and testing.

```{r }
airbnb_manhattan = subset(filter_airbnb, neighbourhood_group == "Manhattan")
airbnb_brooklyn = subset(filter_airbnb, neighbourhood_group == "Brooklyn")

airbnb_manhattan = airbnb_manhattan[,-1]
airbnb_brooklyn = airbnb_brooklyn[,-1]

set.seed(123)
manhattan_sample <- floor(0.8 * nrow(airbnb_manhattan))
manhattan_train_ind <- sample(seq_len(nrow(airbnb_manhattan)), size = manhattan_sample)
manhattan_train <- airbnb_manhattan[manhattan_train_ind, ]
manhattan_test <- airbnb_manhattan[-manhattan_train_ind, ]

brooklyn_sample <- floor(0.8 * nrow(airbnb_brooklyn))
brooklyn_train_ind <- sample(seq_len(nrow(airbnb_brooklyn)), size = brooklyn_sample)
brooklyn_train <- airbnb_brooklyn[brooklyn_train_ind, ]
brooklyn_test <- airbnb_brooklyn[-brooklyn_train_ind, ]

head(brooklyn_train)

```

We want to maximize our net revenue as a lister. Therefore, we want to model net revenue. 
We can control a lot of things such as the listing title, where our property is located,
what kind of property we’re leasing, the price range, and the availability. 
We will create a regression based model to appropriately approach this problem.

```{r }
model_manhattan = lm(formula = log(total_net_revenue) ~ latitude + room_type + 
    calculated_host_listings_count + availability_365 + price_group + usage_group + reviews_per_month, data = manhattan_train)

summary(model_manhattan)

manhattan_predictions = predict(model_manhattan, manhattan_test)
manhattan_predictions = predict(model_manhattan, manhattan_test)

RMSE = function(m, o){sqrt(mean((m - o)^2))}

RMSE(exp(1)^manhattan_predictions, exp(1)^log(manhattan_test$total_net_revenue))
plot(manhattan_predictions, log(manhattan_test$total_net_revenue))
```

After running the regression model on the Manhattan data, we have an Adjusted R-Squared of 0.5478., 
and an RMSE of 17k. So the model is off by about 17k USD / prediction on average.

After running the regression model on the Brooklyn data, we have an Adjusted R-Squared of 0.6263., 
and an RMSE of 23k. So the model is off by about 23k USD / prediction on average.

```{r }
model_brooklyn = lm(formula = log(total_net_revenue) ~ latitude + room_type + price_group + 
    availability_365 + price_group + usage_group + reviews_per_month, data = brooklyn_train)

summary(model_brooklyn)

brooklyn_predictions = predict(model_brooklyn, brooklyn_test)
brooklyn_predictions = predict(model_brooklyn, brooklyn_test)

RMSE(exp(1)^brooklyn_predictions, exp(1)^log(brooklyn_test$total_net_revenue))
plot(brooklyn_predictions, log(brooklyn_test$total_net_revenue))
```

It means that the model is performing better for Manhattan in comparison to Brooklyn
This is a simple baseline model and can be further extended by sing far more complex features 
and advanced machine learning and deep learning algorithms